Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools

### 2024-06-23 paper.qmd
Minor tweaks

  This paper describes an approach for developing strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the underlying methods, including disaggregation of origin-destination data, incorporation of additional trip purposes, routing, scenario generation, and development of an intuitive user interface that has been tested and used by practitioners. Commissioned and used by Transport Infrastructure Ireland, CRUSE provides estimates of current and potential future cycling levels under 'snapshot' scenarios to inform investment decisions. The publicly available results at https://cruse.bike/ enable planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modeling networks at high spatial resolution; simulating multiple trip purposes (social, shopping, personal utility, recreational, and cycle touring) in addition to using official origin-destination data; and providing estimates of 'quietness' (a proxy for cyclist comfort and route preference) at the route segment level. Three network types --- 'Fastest', 'Balanced', and 'Quietest' --- help plan both arterial and residential cycle networks. Workshops and unstructured interviews with stakeholders show that the tool is already being used in practice to inform urban, inter-urban, and rural network planning. The approach is flexible and open-source, allowing the underlying ideas and code to be adapted, supporting effective cycling policies and interventions internationally.

@@

\clearpage

@@

```{r}
#| label: tbl-trip-purposes
#| tbl-cap: "Trip purposes used in the CRUSE Tool."
# For PDF:
# knitr::kable(trip_purposes, booktabs = TRUE) |>
#   kableExtra::kable_styling(latex_options = "striped", full_width = FALSE, font_size = 8) |>
#   kableExtra::column_spec(1, width = "5em") |>
#   kableExtra::column_spec(2, width = "12em") |>
#   kableExtra::column_spec(3, width = "15em") |>
#   kableExtra::column_spec(4, width = "15em")
# For HTML:
knitr::kable(trip_purposes)
```

@@

The CRUSE web application is statically hosted, meaning it does not require a server running software such as the R package 'shiny' or Python packages such as 'streamlit' or 'flask' in the background [@wickham2021].
This reduces hosting costs and the maintenance burden of the tool compared with tools such as the PCT, and makes the map-based interface more responsive, using the recently developed 'MapLibre' JavaScript package and 'PMTiles' "cloud native" vector tile format [@gonçalves2023].
These open source technologies were used to generate default view in the landing page: a map of Ireland with county level results for the % trips that could be made by cycling under the 'Go Dutch' scenario, highlighting the intended use case for strategic cycle network planning and level of ambition, as illustrated in @fig-landing.

@@

## [Stakeholder engagement]{style="color: blue;"} {#sec-stakeholder-engagement}

[The CRUSE Tool was developed for, and in close collaboration with, practitioners in Ireland.
Although the overall approach of the project was quantitative, stakeholder feedback and qualitative data was deemed essential to ensure the tool was fit for purpose, mitigating against the risk that the project would be 'technocentric' and a 'problem looking for a solution'.
From the outset of the project, we aimed to follow @coppola2022's advocacy for a "human-centric approach in which personalization, purpose, ethics and inclusivity are key", rather than an 'off-the-shelf' product that failed to understand, let alone meet, local needs.
As outlined by @coppola2022, the 'digital transformation' taking place in transport planning (and many other sectors) has the potential to increase existing inequalities [@coppola2022].
A wide range of methods and measure can be taken to ensure that digital tools are accessible to their recipients.]{style="color: blue;"}

@@

[The stakeholder engagement was carried out through workshops with stakeholders in case study area, an approach that is documented with reference to participatory development of digital tools for supporting energy transitions [@hewitt2020].
We held a total of three workshops over the course of the project and one open-ended live demo during a 2-day event in Sligo in which we collected feedback from dozens of people.
Other than the live demo, all of the workshops took place online.]{style="color: blue;"}

[The first two workshops took place in May 2022, during which the tool was introduced to key staff involved in cycle network planning in Kildare and Limerick counties.
The county-specific approach allowed the workshop to focus on specific areas and key corridors.
The local transport planners were asked to provide feedback in a semi-structured interview approach, with questions on the evidence they currently have access, how they may be able to use the tool, and requests for new features.
The third and final workshop took place on the 7^th^ December 2023 and included descriptions of recent changes to the tool, with an aim being to get feedback on the new recreational and cycle touring evidence, with representatives from county-level transport planning bodies, Sport Ireland, and Waterways Ireland.]{style="color: blue;"}

>>>>

  This paper describes an approach for developing strategic cycle network planning tools. Based on our experience developing and deploying the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we outline the underlying methods, including disaggregation of origin-destination data with the open source 'odjitter' software, incorporation of additional trip purposes, routing, scenario generation, and development of an intuitive user interface that is tested and used by practitioners. Commissioned by the national infrastructure agency Transport Infrastructure Ireland, CRUSE provides estimates of current and potential future cycling levels under 'snapshot' scenarios to inform investment decisions. The publicly available results at https://cruse.bike/ enable planners, engineers, and other stakeholders to make more evidence-based decisions. CRUSE goes beyond previous work by: modeling networks at high spatial resolution; simulating multiple trip purposes (social, shopping, personal utility, recreational, and cycle touring), supplementing official origin-destination datasets on travel for work and education; and providing estimates of 'quietness' (a proxy for cyclist comfort and route preference) at the route segment level. Three network types --- 'Fastest', 'Balanced', and 'Quietest' --- help plan both arterial and residential cycle networks. Workshops and unstructured interviews with stakeholders were used to inform the development of the tool. Feedback shows that the tool has a wide range of uses and is already being used in practice to inform urban, inter-urban, and rural cycle network designs. The approach is flexible and open-source, allowing the underlying ideas and code to be adapted, supporting more evidence-based and effective cycling policies and interventions internationally.

@@

@@

```{r}
#| label: tbl-trip-purposes
#| tbl-cap: "Trip purposes used in the CRUSE Tool."
# For PDF:
knitr::kable(trip_purposes, booktabs = TRUE) |>
  kableExtra::kable_styling(latex_options = "striped", full_width = FALSE, font_size = 8) |>
  kableExtra::column_spec(1, width = "5em") |>
  kableExtra::column_spec(2, width = "12em") |>
  kableExtra::column_spec(3, width = "15em") |>
  kableExtra::column_spec(4, width = "15em")
# # For HTML:
# knitr::kable(trip_purposes)
```

@@

The CRUSE web application is 'statically hosted' rather than 'dynamically hosted' [@wickham2021].
This means the tool does not require a resource-intensive server, reducing hosting costs and the maintenance burden of the tool compared with tools such as the PCT.
The use recently-developed web mapping technology makes the tool responsive [@gonçalves2023].
Open source technologies were used throughout the tool's user interface, increasing the ease with which people in other countries can build on the open source code.
The 'landing page' shows the proportion of trips that could be made by cycling under the 'Go Dutch' scenario at the county level, highlighting the intended use case for strategic cycle network planning (@fig-landing).

@@

## Stakeholder engagement {#sec-stakeholder-engagement}

The CRUSE Tool was developed for, and in close collaboration with, practitioners in the Republic of Ireland.
Although the overall approach of the project was quantitative, stakeholder feedback and qualitative data was deemed essential to ensure the tool was fit for purpose, mitigating against the risk that the project would be 'technocentric' and a 'problem looking for a solution'.
From the outset of the project, we aimed to follow @coppola2022's advocacy for a "human-centric approach in which personalization, purpose, ethics and inclusivity are key", rather than an 'off-the-shelf' product that failed to understand, let alone meet, local needs.
As outlined by @coppola2022, the 'digital transformation' taking place in transport planning (and many other sectors) has the potential to increase existing inequalities [@coppola2022].
A wide range of methods and measure can be taken to ensure that digital tools are accessible.

@@

The stakeholder engagement was carried out through workshops with stakeholders in case study area, an approach that is documented with reference to participatory development of digital tools for supporting energy transitions [@hewitt2020].
We held a total of three workshops over the course of the project and one open-ended live demo during a 2-day event in Sligo in which we collected feedback from dozens of people.
Other than the live demo, all of the workshops took place online.

The first two workshops took place in May 2022, during which the tool was introduced to key staff involved in cycle network planning in Kildare and Limerick counties.
The county-specific approach allowed the workshop to focus on specific areas and key corridors.
The local transport planners were asked to provide feedback in a semi-structured interview approach, with questions on the evidence they currently have access, how they may be able to use the tool, and requests for new features.
The third and final workshop took place on the 7^th^ December 2023 and included descriptions of recent changes to the tool, with an aim being to get feedback on the new recreational and cycle touring evidence, with representatives from county-level transport planning bodies, Sport Ireland, and Waterways Ireland.

### 2024-06-23 paper.tex
Minor tweaks

\begin{abstract}
This paper describes an approach for developing strategic cycle network
planning tools. Based on our experience developing and deploying the
Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we
outline the underlying methods, including disaggregation of
origin-destination data, incorporation of additional trip purposes,
routing, scenario generation, and development of an intuitive user
interface that has been tested and used by practitioners. Commissioned
and used by Transport Infrastructure Ireland, CRUSE provides estimates
of current and potential future cycling levels under `snapshot'
scenarios to inform investment decisions. The publicly available results
at https://cruse.bike/ enable planners, engineers, and other
stakeholders to make more evidence-based decisions. CRUSE goes beyond
previous work by: modeling networks at high spatial resolution;
simulating multiple trip purposes (social, shopping, personal utility,
recreational, and cycle touring) in addition to using official
origin-destination data; and providing estimates of `quietness' (a proxy
for cyclist comfort and route preference) at the route segment level.
Three network types --- `Fastest', `Balanced', and `Quietest' --- help
plan both arterial and residential cycle networks. Workshops and
unstructured interviews with stakeholders show that the tool is already
being used in practice to inform urban, inter-urban, and rural network
planning. The approach is flexible and open-source, allowing the
underlying ideas and code to be adapted, supporting effective cycling
policies and interventions internationally.
\end{abstract}

@@

\\
\toprule
Approach & Coverage and input data & Outputs and availability & License\\
\midrule
\cellcolor{gray!10}{Bicycle share model, Parkin, J. et al. (2007)} & \cellcolor{gray!10}{England and Wales, Journeys to work OD census data, at the small-area (wards) level} & \cellcolor{gray!10}{Static tables in academic paper} & \cellcolor{gray!10}{Proprietary}\\
Analysis of Cycling Potential (ACP), Transport for London (2010 and 2016) & London, London Travel Demand Survey & Static tables, graphs, and heatmaps, not publicly available & Proprietary\\
\cellcolor{gray!10}{{}[Bikeability index/Bike Score]\{style="color: blue;"\}, [Winters, M. et al. (2013 and 2016)]\{style="color: blue;"\}} & \cellcolor{gray!10}{{}[Metro Vancouver region/160 US and Canadian cities]\{style="color: blue;"\}, [Cycling infrastructure, topography, destinations and road connectivity data]\{style="color: blue;"\}} & \cellcolor{gray!10}{{}[Map-based heatmap, not publicly available]\{style="color: blue;"\}} & \cellcolor{gray!10}{{}[Proprietary]\{style="color: blue;"\}}\\
Prioritization index, Larsen, J. et al. (2013) & Montreal, Survey, Road safety, and OD data & Map-based heatmap, not publicly available & Proprietary\\
\cellcolor{gray!10}{Usage intensity index, Zhang, D. et al. (2014)} & \cellcolor{gray!10}{Belo Horizonte, Survey, census, and OD data} & \cellcolor{gray!10}{Static tables and maps, not publicly available} & \cellcolor{gray!10}{Proprietary}\\
\addlinespace
The Cycling Potential Tool (CPT), Phillips, L., \& Range, A. (2017) & Scotland, Environmental and socioeconomic data, at the small area (output areas) level & Maps showing cycling potential in each area & Proprietary\\
\cellcolor{gray!10}{Propensity to Cycle Tool (PCT), Lovelace, R. et al. (2017)} & \cellcolor{gray!10}{England and Wales, Journeys to work and school OD census data, at the small area (LSOA) level} & \cellcolor{gray!10}{Online maps, graphs, tables, publicly available at www.pct.bike} & \cellcolor{gray!10}{Open source}\\
{}[Cyclist Routing Algorithm for Network Connectivity (CRANC)]\{style="color: blue;"\}, [Gehrke, S. R. et al. (2020)]\{style="color: blue;"\} & {}[Boston region]\{style="color: blue;"\}, [Road segment characteristics, elevation, and destination acessibility data, at the census blocks level]\{style="color: blue;"\} & {}[Graphs and heatmaps, not publicly available]\{style="color: blue;"\} & {}[NA]\{style="color: blue;"\}\\
\cellcolor{gray!10}{The Gross Potential for Cycling tool (CPC), Silva, C. et al. (2021 and 2022) [and Lopez et al. (2021)]\{style="color: blue;"\}} & \cellcolor{gray!10}{21 Portuguese cities, Land use and socio-demographic data, at the small area (census tract) level} & \cellcolor{gray!10}{Static maps showing cycling potential in different areas} & \cellcolor{gray!10}{NA}\\
{}[The relative equity impact of bicycle planning (TIRE)]\{style="color: blue;"\}, [Cunha, I., \& Silva, C. (2023)]\{style="color: blue;"\} & {}[Lisbon]\{style="color: blue;"\}, [Cycling infrastructure, topog­raphy, and location of points of interest data, at the hexagon cells level]\{style="color: blue;"\} & {}[Graphs and heatmaps, not publicly available]\{style="color: blue;"\} & {}[NA]\{style="color: blue;"\}\\
\bottomrule

@@

\clearpage

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.0653}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2869}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4460}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2017}}@{}}

@@

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Trip purpose
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Primary source(s)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Confidence
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Work (commute) & Commuting to / from workplaces & Census 2016 POWSCAR
Data & High confidence - data both origins and destinations for trip
purposes \\
Primary Education & Primary school trips & Census 2016 POWSCAR Data,
Schools database & High confidence - data both origins and destinations
for trip purposes \\
Secondary Education & Secondary school trips & Census 2016 POWSCAR Data,
Schools database & High confidence - data both origins and destinations
for trip purposes \\
Tertiary Education & Tertiary education trips & Census 2016 POWSCAR Data
- Geodirectory categories & High confidence - data both origins and
destinations for trip purposes \\
Social & NHTS classification (e.g.~restaurants, cinemas, gyms etc.) &
Trip rates from NHTS, Geodirectory businesses classified according to
relevant NACE codes & Medium confidence \\
Shopping & NHTS classificationfrom supermarkets and shops & Trip rates
from NHTS, Geodirectory businesses classified according to relevant NACE
codes & Medium confidence \\
Personal / other & Combination of two NHTS classifications, and includes
trips to medical, personal services and others & Trip rates from NHTS,
Geodirectory businesses classified according to relevant NACE codes &
Medium confidence \\
Tourism / recreational & Trips by non-resident visitors to / from
visitor attractions & Geodirectory listings for accommodation (origins)
and destinations (destinations), Failte Ireland accommodation database,
Failte Ireland attraction database & Low confidence - lack of data
regarding tourist trip rates \\

@@

The CRUSE web application is statically hosted, meaning it does not
require a server running software such as the R package `shiny' or
Python packages such as `streamlit' or `flask' in the background
\citep{wickham2021}. This reduces hosting costs and the maintenance
burden of the tool compared with tools such as the PCT, and makes the
map-based interface more responsive, using the recently developed
`MapLibre' JavaScript package and `PMTiles' ``cloud native'' vector tile
format \citep{gonçalves2023}. These open source technologies were used
to generate default view in the landing page: a map of Ireland with
county level results for the \% trips that could be made by cycling
under the `Go Dutch' scenario, highlighting the intended use case for
strategic cycle network planning and level of ambition, as illustrated
in Figure~\ref{fig-landing}.

@@

\subsection{\texorpdfstring{{Stakeholder
engagement}}{Stakeholder engagement}}\label{sec-stakeholder-engagement}

{The CRUSE Tool was developed for, and in close collaboration with,
practitioners in Ireland. Although the overall approach of the project
was quantitative, stakeholder feedback and qualitative data was deemed
essential to ensure the tool was fit for purpose, mitigating against the
risk that the project would be `technocentric' and a `problem looking
for a solution'. From the outset of the project, we aimed to follow
\citet{coppola2022}`s advocacy for a ``human-centric approach in which
personalization, purpose, ethics and inclusivity are key'', rather than
an 'off-the-shelf' product that failed to understand, let alone meet,
local needs. As outlined by \citet{coppola2022}, the `digital
transformation' taking place in transport planning (and many other
sectors) has the potential to increase existing inequalities
\citep{coppola2022}. A wide range of methods and measure can be taken to
ensure that digital tools are accessible to their recipients.}

{The stakeholder engagement was carried out through workshops with
stakeholders in case study area, an approach that is documented with
reference to participatory development of digital tools for supporting
energy transitions \citep{hewitt2020}. We held a total of three
workshops over the course of the project and one open-ended live demo
during a 2-day event in Sligo in which we collected feedback from dozens
of people. Other than the live demo, all of the workshops took place
online.}

{The first two workshops took place in May 2022, during which the tool
was introduced to key staff involved in cycle network planning in
Kildare and Limerick counties. The county-specific approach allowed the
workshop to focus on specific areas and key corridors. The local
transport planners were asked to provide feedback in a semi-structured
interview approach, with questions on the evidence they currently have
access, how they may be able to use the tool, and requests for new
features. The third and final workshop took place on the
7\textsuperscript{th} December 2023 and included descriptions of recent
changes to the tool, with an aim being to get feedback on the new
recreational and cycle touring evidence, with representatives from
county-level transport planning bodies, Sport Ireland, and Waterways
Ireland.}

>>>>

\begin{abstract}
This paper describes an approach for developing strategic cycle network
planning tools. Based on our experience developing and deploying the
Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we
outline the underlying methods, including disaggregation of
origin-destination data with the open source `odjitter' software,
incorporation of additional trip purposes, routing, scenario generation,
and development of an intuitive user interface that is tested and used
by practitioners. Commissioned by the national infrastructure agency
Transport Infrastructure Ireland, CRUSE provides estimates of current
and potential future cycling levels under `snapshot' scenarios to inform
investment decisions. The publicly available results at
https://cruse.bike/ enable planners, engineers, and other stakeholders
to make more evidence-based decisions. CRUSE goes beyond previous work
by: modeling networks at high spatial resolution; simulating multiple
trip purposes (social, shopping, personal utility, recreational, and
cycle touring), supplementing official origin-destination datasets on
travel for work and education; and providing estimates of `quietness' (a
proxy for cyclist comfort and route preference) at the route segment
level. Three network types --- `Fastest', `Balanced', and `Quietest' ---
help plan both arterial and residential cycle networks. Workshops and
unstructured interviews with stakeholders were used to inform the
development of the tool. Feedback shows that the tool has a wide range
of uses and is already being used in practice to inform urban,
inter-urban, and rural cycle network designs. The approach is flexible
and open-source, allowing the underlying ideas and code to be adapted,
supporting more evidence-based and effective cycling policies and
interventions internationally.
\end{abstract}

@@

\\
\toprule
Approach & Coverage and input data & Outputs and availability & License\\
\midrule
\cellcolor{gray!10}{Bicycle share model, Parkin, J. et al. (2007)} & \cellcolor{gray!10}{England and Wales, Journeys to work OD census data, at the small-area (wards) level} & \cellcolor{gray!10}{Static tables in academic paper} & \cellcolor{gray!10}{Proprietary}\\
Analysis of Cycling Potential (ACP), Transport for London (2010 and 2016) & London, London Travel Demand Survey & Static tables, graphs, and heatmaps, not publicly available & Proprietary\\
\cellcolor{gray!10}{Bikeability index/Bike Score, Winters, M. et al. (2013 and 2016)} & \cellcolor{gray!10}{Metro Vancouver region/160 US and Canadian cities, Cycling infrastructure, topography, destinations and road connectivity data} & \cellcolor{gray!10}{Map-based heatmap, not publicly available} & \cellcolor{gray!10}{Proprietary}\\
Prioritization index, Larsen, J. et al. (2013) & Montreal, Survey, Road safety, and OD data & Map-based heatmap, not publicly available & Proprietary\\
\cellcolor{gray!10}{Usage intensity index, Zhang, D. et al. (2014)} & \cellcolor{gray!10}{Belo Horizonte, Survey, census, and OD data} & \cellcolor{gray!10}{Static tables and maps, not publicly available} & \cellcolor{gray!10}{Proprietary}\\
\addlinespace
The Cycling Potential Tool (CPT), Phillips, L., \& Range, A. (2017) & Scotland, Environmental and socioeconomic data, at the small area (output areas) level & Maps showing cycling potential in each area & Proprietary\\
\cellcolor{gray!10}{Propensity to Cycle Tool (PCT), Lovelace, R. et al. (2017)} & \cellcolor{gray!10}{England and Wales, Journeys to work and school OD census data, at the small area (LSOA) level} & \cellcolor{gray!10}{Online maps, graphs, tables, publicly available at www.pct.bike} & \cellcolor{gray!10}{Open source}\\
Cyclist Routing Algorithm for Network Connectivity (CRANC), Gehrke, S. R. et al. (2020) & Boston region, Road segment characteristics, elevation, and destination acessibility data, at the census blocks level & Graphs and heatmaps, not publicly available & NA\\
\cellcolor{gray!10}{The Gross Potential for Cycling tool (CPC), Silva, C. et al. (2021 and 2022) and Lopez et al. (2021)} & \cellcolor{gray!10}{21 Portuguese cities, Land use and socio-demographic data, at the small area (census tract) level} & \cellcolor{gray!10}{Static maps showing cycling potential in different areas} & \cellcolor{gray!10}{NA}\\
The relative equity impact of bicycle planning (TIRE), Cunha, I., \& Silva, C. (2023) & Lisbon, Cycling infrastructure, topog­raphy, and location of points of interest data, at the hexagon cells level & Graphs and heatmaps, not publicly available & NA\\
\bottomrule

@@

\begingroup\fontsize{8}{10}\selectfont

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5em}>{\raggedright\arraybackslash}p{12em}>{\raggedright\arraybackslash}p{15em}>{\raggedright\arraybackslash}p{15em}}

@@

\toprule
Trip purpose & Description & Primary source(s) & Confidence\\
\midrule
\cellcolor{gray!10}{Work (commute)} & \cellcolor{gray!10}{Commuting to / from workplaces} & \cellcolor{gray!10}{Census 2016 POWSCAR Data} & \cellcolor{gray!10}{High confidence - data both origins and destinations for trip purposes}\\
Primary Education & Primary school trips & Census 2016 POWSCAR Data, Schools database & High confidence - data both origins and destinations for trip purposes\\
\cellcolor{gray!10}{Secondary Education} & \cellcolor{gray!10}{Secondary school trips} & \cellcolor{gray!10}{Census 2016 POWSCAR Data, Schools database} & \cellcolor{gray!10}{High confidence - data both origins and destinations for trip purposes}\\
Tertiary Education & Tertiary education trips & Census 2016 POWSCAR Data - Geodirectory categories & High confidence - data both origins and destinations for trip purposes\\
\cellcolor{gray!10}{Social} & \cellcolor{gray!10}{NHTS classification (e.g. restaurants, cinemas, gyms etc.)} & \cellcolor{gray!10}{Trip rates from NHTS, Geodirectory businesses classified according to relevant NACE codes} & \cellcolor{gray!10}{Medium confidence}\\
\addlinespace
Shopping & NHTS classificationfrom supermarkets and shops & Trip rates from NHTS, Geodirectory businesses classified according to relevant NACE codes & Medium confidence\\
\cellcolor{gray!10}{Personal / other} & \cellcolor{gray!10}{Combination of two NHTS classifications, and includes trips to medical, personal services and others} & \cellcolor{gray!10}{Trip rates from NHTS, Geodirectory businesses classified according to relevant NACE codes} & \cellcolor{gray!10}{Medium confidence}\\
Tourism / recreational & Trips by non-resident visitors to / from visitor attractions & Geodirectory listings for accommodation (origins) and destinations (destinations), Failte Ireland accommodation database, Failte Ireland attraction database & Low confidence - lack of data regarding tourist trip rates\\
\bottomrule

@@

The CRUSE web application is `statically hosted' rather than
`dynamically hosted' \citep{wickham2021}. This means the tool does not
require a resource-intensive server, reducing hosting costs and the
maintenance burden of the tool compared with tools such as the PCT. The
use recently-developed web mapping technology makes the tool responsive
\citep{gonçalves2023}. Open source technologies were used throughout the
tool's user interface, increasing the ease with which people in other
countries can build on the open source code. The `landing page' shows
the proportion of trips that could be made by cycling under the `Go
Dutch' scenario at the county level, highlighting the intended use case
for strategic cycle network planning (Figure~\ref{fig-landing}).

@@

\subsection{Stakeholder engagement}\label{sec-stakeholder-engagement}

The CRUSE Tool was developed for, and in close collaboration with,
practitioners in the Republic of Ireland. Although the overall approach
of the project was quantitative, stakeholder feedback and qualitative
data was deemed essential to ensure the tool was fit for purpose,
mitigating against the risk that the project would be `technocentric'
and a `problem looking for a solution'. From the outset of the project,
we aimed to follow \citet{coppola2022}`s advocacy for a ``human-centric
approach in which personalization, purpose, ethics and inclusivity are
key'', rather than an 'off-the-shelf' product that failed to understand,
let alone meet, local needs. As outlined by \citet{coppola2022}, the
`digital transformation' taking place in transport planning (and many
other sectors) has the potential to increase existing inequalities
\citep{coppola2022}. A wide range of methods and measure can be taken to
ensure that digital tools are accessible.

The stakeholder engagement was carried out through workshops with
stakeholders in case study area, an approach that is documented with
reference to participatory development of digital tools for supporting
energy transitions \citep{hewitt2020}. We held a total of three
workshops over the course of the project and one open-ended live demo
during a 2-day event in Sligo in which we collected feedback from dozens
of people. Other than the live demo, all of the workshops took place
online.

The first two workshops took place in May 2022, during which the tool
was introduced to key staff involved in cycle network planning in
Kildare and Limerick counties. The county-specific approach allowed the
workshop to focus on specific areas and key corridors. The local
transport planners were asked to provide feedback in a semi-structured
interview approach, with questions on the evidence they currently have
access, how they may be able to use the tool, and requests for new
features. The third and final workshop took place on the
7\textsuperscript{th} December 2023 and included descriptions of recent
changes to the tool, with an aim being to get feedback on the new
recreational and cycle touring evidence, with representatives from
county-level transport planning bodies, Sport Ireland, and Waterways
Ireland.
